---
title: "Working With GOES AODC Rasters"
author: "Mazama Science"
date: "8/12/2019"
output: html_document
vignette: >
  %\VignetteIndexEntry{Working With GOES AODC Rasters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MazamaSatelliteUtils)
library(MazamaSpatialUtils)
setSatelliteDataDir("~/Data/Satellite")
setSpatialDataDir("~/Data/Spatial")
loadSpatialData("USCensusStates")

florida <- subset(USCensusStates, stateCode == "FL")
nc <- goesaodc_openFile("OR_ABI-L2-AODC-M6_G16_s20191291201274_e20191291204047_c20191291210009.nc")

pal_aod <- colorRampPalette(c("lightgoldenrod1", "red3"))
pal_dqf <- colorRampPalette(c("green", "red"))
```

## Raster Objects

A Raster is used to store and visualize information that is "gridded" like the
pixels of an image. It is tailored towards visualizing geospatial data so that
each pixel (or "cell") of the Raster covers a small area of the Earth's 
surface and the color represents the value of a measurement, such as aerosol 
optical depth. A Raster is created in several steps, but the process can be 
streamlined with `MazamaSatelliteUtils` functions. Let's walk through the 
process first though to understand what Rasters are made of and how they are put
together:

[more](https://www.neonscience.org/raster-res-extent-pixels-r)

## Creating a RasterLayer

Before any data can actually be rastered, a RasterLayer must first be defined in
order to to set up the grid to draw to. This grid is given a bounding box with
min/max longitudes and latitudes as well as a horizontal and vertical resolution
given the number of rows and columns of cells. Like a digital image, more pixels
results in a sharper and less blocky result.

The last requirement is a Coordinate Reference System (CRS) `proj4string`. This 
is necessary to project the RasterLayer over the correct area on Earth using
standard points of reference.
"+proj=longlat +datum=WGS84 +ellps=GRS80" is the default we use and can be 
broken up into several elements:

1. proj=longlat: The projection system to use in order to transform 3D spherical
world coordinates onto our 2D screen. In this case we'll use longitude and
latitude. There are quite a lot of 
[others](https://proj.org/operations/projections/index.html) to choose from.

2. datum=WGS84: A coordinate system for the Earth. The WGS84 (World Geodetic 
System 1984) is a common standard and is used for Google Maps.

3. ellps=GRS80: The ellipsoid used to model the Earth. GRS80 and WGS84 are
almost identical.

Further reading on CRS and projections:
[More on geodetic systems](https://gisgeography.com/wgs84-world-geodetic-system/)
[PROJ4](https://proj.org/)
[more on CRS](https://rspatial.org/spatial/6-crs.html).
[Overview of Coordinate Reference Systems (CRS) in R](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf)
[Coordinate Systems in R](https://mgimond.github.io/Spatial/coordinate-systems-in-r.html)

```{r}
res <- 0.05
bbox <- sp::bbox(florida)

lon_min <- bbox[1, 1]
lon_max <- bbox[1, 2]
lat_min <- bbox[2, 1]
lat_max <- bbox[2, 2]

ncols <- ((lon_max - lon_min) / res) + 1
nrows <- ((lat_max - lat_min) / res) + 1

rasterLayer <- raster::raster(nrows = nrows, ncols = ncols,
                              xmn = lon_min, xmx = lon_max,
                              ymn = lat_min, ymx = lat_max,
                              crs = "+proj=longlat +datum=WGS84 +ellps=GRS80")
```

## Rasterizing SpatialPoints

Once the `RasterLayer` grid is defined, we can rasterize all of the readings 
from our `SpatialPoints` object with the `rasterize` function. Now we have to 
decide how our cells should be colored based on SpatialPoint measurements. A 
cell will not be defined if it doesn't contain any points, but what if it 
contains more than one point? For this case we define the `fun` parameter for 
`rasterize`, which determines how to calculate one summarizing value from many
points. By default this is set to be the "mean" value for all the points in the 
cell.

The result of this function is a `RasterBrick` with a separate `RasterLayer` for 
each variable.

```{r, echo=FALSE}
spatialPoints <- goesaodc_createSpatialPoints(nc = nc, bbox = bbox)
rasterBrick <- raster::rasterize(spatialPoints, rasterLayer, fun = mean)
plot(rasterBrick)
```

The "ID" layer is included automatically--coloring each cell by the ID numbers
of the features in the `SpatialPoints` object. While it's not very helpful in
determining air quality, it does give a nice visual for how the GOES-16 ABI
scans the Earth from North to South. As the ID plot legend suggests, there are
about 70,000 SpatialPoint features in all:

```{r}
spatialPoints
```

The AOD and DQF measurements are more useful for our atmospheric analysis. Let's
plot them with the state border on top:

```{r, echo=FALSE, fig.width=8, fig.height=4}
par(mfrow=c(1, 2), mar=c(2, 4, 2, 4))

plot(rasterBrick$AOD, main = "AOD", col = pal_aod(50))
plot(florida, add = TRUE)

plot(rasterBrick$DQF, main = "DQF", col = pal_dqf(50))
plot(florida, add = TRUE)
```

The process for creating this final RasterBrick is streamlined by the 
`MazmaSatelliteUtils` function `goesaodc_createRaster`. The following code 
produces the exact same result without having to go through the steps of 
explicitly defining a new RasterLayer and rastering to it:

```{r}
rb_simpler <- goesaodc_createRaster(nc = nc, res = res, fun = mean, bbox = bbox)

plot(rb_simpler$AOD, main = "AOD", col = pal_aod(50))
plot(florida, add = TRUE)
```

## Raster Metrics

We can select several different `fun` methods besides `mean` to calculate the 
final value for a cell. There are the usual statistical functions: `median`, 
`sd`, `min`, `max`, as well as several specific functions provided by rasterize:
"count", "first", and "last". "last" is used by default.

```{r, fig.width=8, fig.height=6}
rb_mean   <- raster::rasterize(spatialPoints, rasterLayer, fun = mean)
rb_median <- raster::rasterize(spatialPoints, rasterLayer, fun = median)
rb_sd     <- raster::rasterize(spatialPoints, rasterLayer, fun = sd)
rb_count  <- raster::rasterize(spatialPoints, rasterLayer, fun = "count")

par(mfrow=c(2, 2), mar=c(2, 3, 2, 3))
plot(rb_mean$AOD,   main = "AOD Mean", col = pal_aod(50))
plot(rb_median$AOD, main = "AOD Median", col = pal_aod(50))
plot(rb_sd$AOD,     main = "AOD Standard Deviation", col = pal_aod(50))
plot(rb_count$AOD,  main = "Points in Cells", col = pal_aod(50))
```

## RasterStacks and Averaging

```{r}
rasterVis::levelplot(rasterBrick$AOD, main = "AOD", margin = FALSE)
```