---
title: "Working with AODC Data Structures"
author: "Mazama Science"
date: "1/8/2020"
output: html_document
vignette: >
  %\VignetteIndexEntry{Working with AODC Data Structures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MazamaSatelliteUtils)
library(MazamaSpatialUtils)
setSatelliteDataDir("~/Data/Satellite")
setSpatialDataDir("~/Data/Spatial")
loadSpatialData("USCensusStates")
```

The **MazamaSpatialUtils** package utilizes several different data structures that 
represent different levels of processing of GOES AOD data. As stored in the
NetCDF files, the AOD is on its "native" pixel grid, with `x` and `y` representing
scan angles as opposed to longitude and latitude. Each image is a snapshot of
the spherical earth and thus the grid is inherently curvilinear.

This raw, curvilinear data is processed into different data structures, step-by-step
with the following functions:

1. `createNativeGrid()` -- returns `list` of `array`s with native pixel grid
2. `createTibble()` -- returns `tibble` with longitude, latitude, value
3. `createSpatialPoints()` -- returns **sp** package `SpatialPointsDataFrame`
4. `createRaster()` -- returns **raster** package `raster`
5. `createRasterStack()` -- returns **raster** package `rasterStack`

Each successive data structure is described in the sections below.

Here we set things up with data from the 2019 Kincade fire in California:

```{r kincade_setup}
library(MazamaSatelliteUtils)

setSatelliteDataDir("~/Data/Satellite")

goesaodc_downloadAOD(
  satID = "G17",
  datetime = "2019-10-27 14:00",
  timezone = "America/Los_Angeles"
)

files <- goesaodc_listFiles("G17", "2019-10-27 14:00", timezone = "America/Los_Angeles")

ncList <- list()
for ( file in files ) {
  label <-
    file %>%
    goesaodc_convertFilenameToDatetime() %>%
    MazamaCoreUtils::timeStamp(unit = "sec", timezone = "UTC")
  ncList[[label]] <- goesaodc_openFile(basename(file))
}
```

## NativeGrid

At the lowest level, data are accessible on the native, curvilinear grid by 
using the `goesaodc_nativGrid()` function. This function finds the native grid
cells which are bounded by the incoming `bbox` and returns a `list` of `array`s
named `lon`, `lat`, `AOD` and `DQF`. Individual elements can be viewd with
`graphics::image()`.

While GOES AODC CONUS images are commonly 3+ Megabytes, specification of `bbox`
allows ingest of much smaller subsets with a minimal memory footprint.

_Note that we have to reverse the j indices when using `base::image()`._

```{r native_grid, fig.width = 6, fig.height = 6}
# Using previously defined variables
kincade_bbox <- c(-124, -120, 36, 39)

nativeGrid <- goesaodc_createNativeGrid(ncList[[1]], kincade_bbox)

j_indices <- ncol(nativeGrid$lon):1

layout(matrix(1:4, nrow = 2))

image(nativeGrid$lon[, j_indices])
title("Native Grid -- lon")
image(nativeGrid$lat[, j_indices])
title("Native Grid -- lat")
image(nativeGrid$AOD[, j_indices])
title("Native Grid -- AOD")
image(nativeGrid$DQF[, j_indices])
title("Native Grid -- DQF")

layout(1)
```
    
## Tibble
  
Once the desired data have been ingested, the next step is to create a `tibble`
(modern `data.frame`) where each record has the `lon, lat, AOD, DQF` associated
with a single pixel. Internally, this data structure is used to subset data
based on rectangular lat-lon boundaries.

This datas tructure and the _nativeGrid_ list are primarily used internally and
rarely of interest to an end user. The `goesaodc_createTibble()` calls
`goesaodc_createNativeGrid() internally.

```{r tibble}
# Using previously defined variables
tibble <- goesaodc_createTibble(ncList[[1]], kincade_bbox)
head(tibble)
```
    
## Spatial Points
  
The first _spatial_ data structure is created by `goesaodc_createSpatialPoints()`
which calls `goesaodc_createTibble()` internally and uses the **sp** package to
convert the tibble into a `SpatialPointsDataFrame`. This data structure is 
useful for creating maps showing pixels in their native grid position.

The `goesaodc_plotSpatialPoints()` function can plot the AOD values from this 
data structure as filtered by the requested `dqfLevel`.
    
```{r spatialPoints}
# Using previously defined variables
SPDF <- goesaodc_createSpatialPoints(ncList[[1]], dqfLevel = 3, bbox = kincade_bbox)
maps::map(database = "state", "regions" = c("california"), xlim = c(-126, -113.5))
goesaodc_plotSpatialPoints(SPDF, cex = 0.2, add = TRUE)
title("SpatialPoints AOD (DQF <= 3)")
```

## Raster
  
For basic plotting needs, the `SpatialPointsDataFrame` is the highest data 
structure needed. But the **raster** package has many advanced capabilities for
image processing once data have been converted into a `raster` -- a data
structure with a _rectilinear_ rather than a _curvilinear_ structure in
longitude-latitude space.

The `goesaodc_createRaster()` function takes care of loading the requested 
data and assigning data to a _rectilinear_ grid determined by the `res`
argument specifying degrees per grid cell. When multiple valid pixel values
occur within a grid cell, the grid cell value is determined by applying the
function specified with `fun` (averaging by default).

This function should be used when working with larger areas where some spatial
averaging is desired. It is also possible to create rasters whose underlying
grid matches that of a particular model run -- allowing direct comparison.

```{r raster}
# Using previously defined variables
ca_nv <- subset(USCensusStates, stateCode %in% c("CA","NV"))
ca_nv_bbox <- bbox(ca_nv)

rstr <- goesaodc_createRaster(
  ncList[[1]],
  res = 0.05,
  fun = mean,
  dqfLevel = 3,
  bbox = ca_nv_bbox
)

raster::plot(rstr, "AOD", col = rev(heat.colors(5)), axes = FALSE)
plot(ca_nv, add = TRUE)
title("Raster -- AOD (DQF <= 3)")
```
    
## RasterStack
  
At the highest level of data processing is the **raster** package `rasterStack` 
data structure which combines multiple `raster` objects, typically associated
with individual time steps.

The **rasterVis** package provides convenient functions for creating data
visualizations covering multiple timesteps as seen in the example below.

The `goesaodc_createRasterStack()` is currently an experimental function that
does has not been fuly harmonized with other package functions. **It will 
change in a future version of the package.**

As currently implemented, the `goesaodc_createRasterStack()` does not take a
vector of `nc` file handles. As currently implemented -- _this may change_ --
the function is designed to accept information on the desired satellite,
time range and area and takes care of finding and potentially downloading 
data files, opening NetCDF files, ingesting data and then properly closing the
NetCDF files.

It is also possible to specify a vector of file names but the `satID` and
`datetime` arguments must still be specified. Again, _this may change_ in a
future version


```{r rasterStack}
# Start by closing the nc file handles we opened at the beginning
for ( nc in ncList ) {
  ncdf4::nc_close(nc)
}

# Using previously defined variables
rasterStack <- goesaodc_createRasterStack(
  satID = "G17",
  datetime = "2019-10-27 14:00",
  endTime = "2019-10-27 15:00",
  bbox = kincade_bbox,
  dqfLevel = 3,
  timezone = "America/Los_Angeles",
  res = 0.075,
  isJulian = FALSE,
  fileList = files,
  verbose = FALSE
)

rasterVis::levelplot(rasterStack, par.settings = rasterVis::YlOrRdTheme())
```
    
Best of luck working with AOD satellite data!
