---
title: "Grid Sizes"
author: "Mazama Science"
date: "8/2/2019"
output: html_document
vignette: >
  %\VignetteIndexEntry{Grid Sizes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MazamaSatelliteUtils)
library(MazamaSpatialUtils)

setSatelliteDataDir("~/Data/Satellite")
setSpatialDataDir("~/Data/Spatial")

# get bbox for Texas
loadSpatialData("USCensusStates")
tx <- subset(USCensusStates, stateCode == "TX")
bbox_tx <- sp::bbox(tx)

nc <- goesaodc_openFile("OR_ABI-L2-AODC-M6_G16_s20191291201274_e20191291204047_c20191291210009.nc")
pal_aod <- colorRampPalette(c("lightgoldenrod1", "red3"))
pal_count <- colorRampPalette(c("aliceblue", "dodgerblue4"))
pal_sd <- colorRampPalette(c("lightpink", "darkred"))
```

Creating a raster image divides geographic space into uniform pixels or "cells".
The resolution of the image defines how much world space each cell represents, 
so larger cells cover a larger area and (usually) more data points. A single 
cell is created for a region whether it contains hundreds of data points or just
one. This raises two issues:

1. When a cell is sparsely populated, it will define the entire region it covers
by only a few values. This will not accurately represent the whole area.

2. When a cell is densly populated, the it will have to generalize all of the 
data in order to determine its single color value. This may not correctly depict
potential disparities in AOD throughout the area.

In this vignette we will be analyzing raster grids for these different 
scenarios. A good place to find areas of both types is Texas on ?????, where the 
sea just off the east coast is dense with data points as opposed to the more 
sparse center of the state:

```{r plot_texas_points, fig.width=5, fig.height=5, fig.align='center', echo=FALSE}
tb <- goesaodc_createTibble(nc)
data_tx <- dplyr::filter(tb, 
                         lon > bbox_tx["x", "min"], lon < bbox_tx["x", "max"], 
                         lat > bbox_tx["y", "min"], lat < bbox_tx["y", "max"])

sp <- goesaodc_createSpatialPoints(nc, bbox = bbox_tx)

par(mar=c(0, 0, 2, 0))
plot(tx, main = "Texas AOD")
goesaodc_plotSpatialPoints(sp, var = "AOD", cex = 0.25, add = TRUE)
plot(tx, add = TRUE)
```

Here are three rasters with different resolutions. Notice how the lower 
resolution grids suggest there is data for areas where no AOD values were even 
measured:

```{r texas_resolution_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
par(mfrow=c(1, 3), mar=c(2, 1, 2, 1))
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.1)
raster::plot(rstr$AOD, main = "AOD Res. 0.1", col = pal_aod(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.3)
raster::plot(rstr$AOD, main = "AOD Res. 0.3", col = pal_aod(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 1.0)
raster::plot(rstr$AOD, main = "AOD Res. 1.0", col = pal_aod(100), legend = FALSE)
plot(tx, add = TRUE)
```

As the tiles increase in size they become more and more inaccurate in 
summarizing the true aerosol depth over the area they cover. Let's start 
analyzing the two cases from earlier though, starting with the very dense cloud 
of points off the southeastern coast near Corpus Christi:

```{r dense_plot, fig.width=8, fig.height=6, fig.align='center', echo=FALSE}
# Subset for the east coast of Texas where there is a lot of dense data
bb_tx_dense <- c(-97, -96, 27, 28)
tb_tx_dense <- dplyr::filter(tb, 
                             lon > bb_tx_dense[1], lon < bb_tx_dense[2], 
                             lat > bb_tx_dense[3], lat < bb_tx_dense[4])
rstr_mean <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = mean)

# Plot the raster colored with the AOD palette and draw all sample point locations
raster::plot(rstr_mean$AOD, main = "AOD", col = pal_aod(25), xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, pch = 16, cex = 0.4)
```

This plot helps illustrate how the GOES 16 satellite arranges its sample
locations. Even in this dense area though there are still gaps and holes in the
readings. The cell colors in this plot are given by the mean AOD value of all
the points within the area they cover. In the more uniform regions this may 
average around twenty points, but in the more spotty regions in the upper-left
we can see that some tiles are defined by less than five values.

Drawing the sample points with the same palette shows how the raster
generalizes the actual AOD readings within each cell. Let's compare how 
different our results are when using either the mean or median: 

```{r dense_average_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
# Assign each reading its own color based on the AOD palette
tb_tx_dense$col <- pal_aod(25)[as.numeric(cut(tb_tx_dense$AOD, breaks = 25))]

# Generate the median raster
rstr_median <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = median)

par(mfrow = c(1, 2), mar = c(2, 4.5, 2, 4.5))

# Plot the mean raster
raster::plot(rstr_mean$AOD, main = "Mean AOD", col = pal_aod(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, 
       pch = 16, cex = 0.5, col = tb_tx_dense$col)

# Plot the median raster
raster::plot(rstr_median$AOD, main = "Median AOD", col = pal_aod(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, 
       pch = 16, cex = 0.5, col = tb_tx_dense$col)
```

The two raster images turn out to be pretty similar. One could argue though that
using the median is a better representation though since the cells of that map
match the colors of their points more unformly so you can't even tell where they
are.

We can color the cells using different summarizing statistics besides the
mean and median. Rasters can be generated based on the minimum data value in a 
cell, the maximum value, the first or last, the total number of values, the 
standard deviation, and several others. Let's try out the count and standard 
deviation options:

```{r dense_extra_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
rstr_count <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = "count")

rstr_sd <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = sd)

par(mfrow = c(1, 2), mar = c(2, 4.5, 2, 4.5))
raster::plot(rstr_count$AOD, main = "Point Count", col = pal_count(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, pch = 16, cex = 0.3)


raster::plot(rstr_sd$AOD, main = "AOD Standard Deviation", col = pal_sd(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE, main = "AOD Standard Deviation")
points(tb_tx_dense$lon, tb_tx_dense$lat, pch = 16, cex = 0.3)
```

Now that we know what is going on in a small area, let's zoom back out to the 
state overview and see how various resolutions influence the apparent point
counts and standard deviations:

```{r texas_count_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
par(mfrow=c(1, 3), mar=c(2, 1, 2, 1))
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.1, fun = "count")
raster::plot(rstr$AOD, main = "Point Count Res: 0.1", col = pal_count(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.3, fun = "count")
raster::plot(rstr$AOD, main = "Point Count Res: 0.3", col = pal_count(100), legend = FALSE)
plot(tx, add = TRUE)
  
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 1.0, fun = "count")
raster::plot(rstr$AOD, main = "Point Count Res: 1.0", col = pal_count(100), legend = FALSE)
plot(tx, add = TRUE)
```

```{r texas_sd_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
par(mfrow=c(1, 3), mar=c(2, 1, 2, 1))
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.1, fun = sd)
raster::plot(rstr$AOD, main = "AOD SD Res: 0.1", col = pal_sd(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.3, fun = sd)
raster::plot(rstr$AOD, main = "AOD SD Res: 0.3", col = pal_sd(100), legend = FALSE)
plot(tx, add = TRUE)
  
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 1.0, fun = sd)
raster::plot(rstr$AOD, main = "AOD SD Res: 1.0", col = pal_sd(100), legend = FALSE)
plot(tx, add = TRUE)
```