---
title: "Grid Sizes"
author: "Mazama Science"
date: "8/2/2019"
output: html_document
vignette: >
  %\VignetteIndexEntry{Grid Sizes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MazamaSatelliteUtils)
library(MazamaSpatialUtils)

setSatelliteDataDir("~/Data/Satellite")
setSpatialDataDir("~/Data/Spatial")

# get bbox for Texas
loadSpatialData("USCensusStates")
tx <- subset(USCensusStates, stateCode == "TX")
bbox_tx <- sp::bbox(tx)

nc <- goesaodc_openFile("OR_ABI-L2-AODC-M6_G16_s20191291201274_e20191291204047_c20191291210009.nc")
pal_aod <- colorRampPalette(c("lightgoldenrod1", "red3"))
pal_count <- colorRampPalette(c("aliceblue", "dodgerblue4"))
pal_sd <- colorRampPalette(c("lightpink", "darkred"))
```

A raster image divides geographic space into uniform pixels or "cells".
The uniform size of these cells is determined by the resolution of the raster so that
higher resolutions reduce cell area and increases the detail and accuracy of
the final image. The `goesaodc_createRaster()` function defines a colored 
cell for each grid space with at least one data point, which raises two issues:

1. When a grid space is sparsely populated, a cell will be defined for the whole
area using only a few data possibly concentrated in just a small corner. These
might not accurately represent the AOD level for the rest of the cell.

2. When a grid space is densely populated, the cell will have to generalize a 
lot of data values in order to determine its final color. If there is a 
large disparity in AOD throughout the region, this will not be visible if the
cell color is based on the average or median reading level.

In this vignette we will be analyzing raster grids for both
scenarios. A good place to find areas of both types is Texas on ?????, where the 
sea just off the east coast is dense with data points as opposed to the more 
sparse center of the state:

```{r plot_texas_points, fig.width=5, fig.height=5, fig.align='center', echo=FALSE}
tb <- goesaodc_createTibble(nc)
data_tx <- dplyr::filter(tb, 
                         lon > bbox_tx["x", "min"], lon < bbox_tx["x", "max"], 
                         lat > bbox_tx["y", "min"], lat < bbox_tx["y", "max"])

sp <- goesaodc_createSpatialPoints(nc, bbox = bbox_tx)

par(mar=c(0, 0, 2, 0))
plot(tx, main = "Texas AOD")
goesaodc_plotSpatialPoints(sp, var = "AOD", cex = 0.15, add = TRUE)
plot(tx, add = TRUE)
```

Here are three rasters with different resolutions. Notice how the lower 
resolution grids suggest there is data for areas where no AOD measurements have
been taken at all:

```{r texas_resolution_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
par(mfrow=c(1, 3), mar=c(2, 1, 2, 1))
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.1)
raster::plot(rstr$AOD, main = "AOD Res. 0.1", col = pal_aod(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.3)
raster::plot(rstr$AOD, main = "AOD Res. 0.3", col = pal_aod(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 1.0)
raster::plot(rstr$AOD, main = "AOD Res. 1.0", col = pal_aod(100), legend = FALSE)
plot(tx, add = TRUE)
```

As the tiles increase in size they become more and more inaccurate in 
summarizing the true aerosol depth over the area they cover. Let's zoom to the 
dense cloud of points off the southeastern coast near Corpus Christi so we can
better undsertand how data points influence individual cells. Each measurement location
is marked with a black dot: 

```{r dense_plot, fig.width=8, fig.height=6, fig.align='center', echo=FALSE}
# Subset for the east coast of Texas where there is a lot of dense data
bb_tx_dense <- c(-97, -96, 27, 28)
tb_tx_dense <- dplyr::filter(tb, 
                             lon > bb_tx_dense[1], lon < bb_tx_dense[2], 
                             lat > bb_tx_dense[3], lat < bb_tx_dense[4])
rstr_mean <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = mean)

# Plot the raster colored with the AOD palette and draw all sample point locations
raster::plot(rstr_mean$AOD, main = "AOD", col = pal_aod(25), xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, pch = 16, cex = 0.4)
```

Here we can see how the GOES-16 satellite takes its sample measurements in a
non-orthogonal grid. Even in this uniform arrangement though there are still 
holes and gaps left in the readings. The cell colors in this plot are given by 
the mean AOD value of all the points within the area they cover. In the dense 
spaces of the grid this may end up averaging around twenty points while the more
more sparse regions to the upper-left may only use five or fewer values.

Now drawing the sample points with the same AOD palette shows how the raster
generalizes the individual readings within each cell. Let's compare how 
different our results are when using either the mean or median: 

```{r dense_average_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
# Assign each reading its own color based on the AOD palette
tb_tx_dense$col <- pal_aod(25)[as.numeric(cut(tb_tx_dense$AOD, breaks = 25))]

# Generate the median raster
rstr_median <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = median)

par(mfrow = c(1, 2), mar = c(2, 4.5, 2, 4.5))

# Plot the mean raster
raster::plot(rstr_mean$AOD, main = "Mean AOD", col = pal_aod(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, 
       pch = 16, cex = 0.5, col = tb_tx_dense$col)

# Plot the median raster
raster::plot(rstr_median$AOD, main = "Median AOD", col = pal_aod(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, 
       pch = 16, cex = 0.5, col = tb_tx_dense$col)
```

The two raster images turn out to be pretty similar. One could argue though that
the median gives a better representation since the cells of that map blend better
with the colors of all their points for some regions.

It's possible to use different summarizing statistics besides the mean and 
median to color cells. Rasters can be generated based on the minimum data value in a 
cell, the maximum value, the first or last, the total number of values, the 
standard deviation, and several other ways. Let's try out the count and standard 
deviation options:

```{r dense_extra_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
rstr_count <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = "count")

rstr_sd <- goesaodc_createRaster(nc, 
                              lonLo = bb_tx_dense[1], lonHi = bb_tx_dense[2], 
                              latLo = bb_tx_dense[3], latHi = bb_tx_dense[4], 
                              res = 0.1, fun = sd)

par(mfrow = c(1, 2), mar = c(2, 4.5, 2, 4.5))
raster::plot(rstr_count$AOD, main = "Point Count", col = pal_count(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE)
points(tb_tx_dense$lon, tb_tx_dense$lat, pch = 16, cex = 0.3)


raster::plot(rstr_sd$AOD, main = "AOD Standard Deviation", col = pal_sd(25), 
             xlim = bb_tx_dense[1:2], ylim = bb_tx_dense[3:4])
plot(tx, add = TRUE, main = "AOD Standard Deviation")
points(tb_tx_dense$lon, tb_tx_dense$lat, pch = 16, cex = 0.3)
```

Due to the irregularity of the data point layout, the count plot displays a
sort of ripple effect. Meanwhile, the standard deviation plot is useful to determine
areas where we may encounter the problem from earlier where a cell has to generalize 
values that differ greatly.

Now that we know what is going on in a small area, let's zoom back out to the 
state overview and see how various resolutions influence the apparent point
counts and standard deviations:

```{r texas_count_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
par(mfrow=c(1, 3), mar=c(2, 1, 2, 1))
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.1, fun = "count")
raster::plot(rstr$AOD, main = "Point Count Res: 0.1", col = pal_count(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.3, fun = "count")
raster::plot(rstr$AOD, main = "Point Count Res: 0.3", col = pal_count(100), legend = FALSE)
plot(tx, add = TRUE)
  
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 1.0, fun = "count")
raster::plot(rstr$AOD, main = "Point Count Res: 1.0", col = pal_count(100), legend = FALSE)
plot(tx, add = TRUE)
```

```{r texas_sd_plots, fig.width=10, fig.height=4, fig.align='center', echo=FALSE}
par(mfrow=c(1, 3), mar=c(2, 1, 2, 1))
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.1, fun = sd)
raster::plot(rstr$AOD, main = "AOD SD Res: 0.1", col = pal_sd(100), legend = FALSE)
plot(tx, add = TRUE)

rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 0.3, fun = sd)
raster::plot(rstr$AOD, main = "AOD SD Res: 0.3", col = pal_sd(100), legend = FALSE)
plot(tx, add = TRUE)
  
rstr <- goesaodc_createRaster(nc, bbox = bbox_tx, res = 1.0, fun = sd)
raster::plot(rstr$AOD, main = "AOD SD Res: 1.0", col = pal_sd(100), legend = FALSE)
plot(tx, add = TRUE)
```