---
title: "Comparing G16 and G17"
author: "Mazama Science"
date: "5/20/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing G16 and G17}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Although parts of the GOES-16/17 satellite scan ranges overlap (See 
'GOES Coverage' article), the scans they produce can turn out quite different.
For example, take a look at their scans for the same moment over the Cherrywood
Fire in Nevada (about (37, -116)).

```{r, echo=FALSE, fig.width=7, fig.height=5, fig.align="center"}
library(MazamaSatelliteUtils)
library(MazamaSpatialUtils)

setSatelliteDataDir("~/Data/Satellite")
setSpatialDataDir("~/Data/Spatial")

loadSpatialData("NaturalEarthAdm1")

bbox <- c(-122.5, -111.5, 34.5, 42.5)

scanFileG16 <- goesaodc_downloadScanFiles(
  satID = "G16",
  datetime = "2021-05-20 00:30",
  timezone = "UTC"
)

scanRasterG16 <- goesaodc_createScanRaster(
  filename = scanFileG16,
  bbox = bbox,
  dqfLevel = 3,
  cellSize = 0.05
)

goesaodc_plotScanRaster(
  raster = scanRasterG16,
  bbox = bbox,
  legendLimits = c(-0.1, 5.1),
  stateCodes = c("OR", "ID", "NV", "CA", "AZ", "UT"),
  title = "G16 at 2021-05-20 00:30 UTC"
)
```

```{r, echo=FALSE, fig.width=7, fig.height=5, fig.align="center"}
scanFileG17 <- goesaodc_downloadScanFiles(
  satID = "G17",
  datetime = "2021-05-20 00:30",
  timezone = "UTC"
)

scanRasterG17 <- goesaodc_createScanRaster(
  filename = scanFileG17,
  bbox = bbox,
  dqfLevel = 3,
  cellSize = 0.05
)

goesaodc_plotScanRaster(
  raster = scanRasterG17,
  bbox = bbox,
  legendLimits = c(-0.1, 5.1),
  stateCodes = c("OR", "ID", "NV", "CA", "AZ", "UT"),
  title = "G17 at 2021-05-20 00:30 UTC"
)
```

There are several differences between these two images. For instance, GOES-17 
appears to record more cloud cover than GOES-16. Additionally, GOES-16 reads AOD
values over the ocean (maybe all bodies of water?) as ~0, while GOES-17 reads 
them as ~5.

Even though GOES-17 is positioned closer to Nevada, from these two plots we can
see that the clearest images we'd get of the Cherrywood fire would come from
GOES-16 scans. A bit counterintuitive, isn't it?