---
title: "Visualizing GOES AOD"
author: "Mazama Science"
date: "4/21/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing GOES AOD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This article will walk through how to visualize GOES AOD scans using high-level 
functions available in the **MazamaSatelliteUtils** package.

## Point and Raster Plots 

Each GOES scan records the AOD value for millions of fixed locations across the 
continental US. **MazamaSatelliteUtils** allows you to visualize this data 
either as points or a raster--each having their upsides and downsides when it 
comes to plotting.

### Point Plots

**Pros:**

- Points show the raw (unaggregated) AOD values at the exact location where they 
were sampled. Therefore, points give the most direct information at the highest 
possible spatial resolution.
- Render very fast in both Mercator-projected and Cartesian coordinate systems.

**Cons:**

- The arrangement of sample locations means points will likely overlap or have 
gaps between them when plotted.
- Point size is not determined by axis units. Therefore, the user will have to 
manually determine the best point size for each plot they make depending on the
plot dimensions and the area it covers.

### Raster Plots

**Pros:**

- Cells aggregate multiple points in a small region and therefore help fill 
'gaps' of `NA` AOD values.
- Cells will never overlap of have gaps between them and their size is defined 
using axis units. Therefore, the user does not have to adjust anything 
depending on the size of the plot or the area it covers.

**Cons:**

- Cells will not be as precise (both in raw AOD value and location) as points,
unless the grid resolution is _very_ high.
- High-resolution rasters render slowly in Mercator-projected coordinate 
systems.

## How to Plot a Scan

Visualizing GOES AOD data can be done in just a few steps:

### 1. Getting Scan Files

First things first: since each scan is saved as a netCDF file on a remote 
server, we need to download the files for the scans we want to see! We can do
this with `goesaodc_downloadScanFiles()`, which downloads the scan file taken 
closest to the given `datetime` by the named satellite. If an `endtime` is given
as well, then all the satellite's scan files taken at or after `datetime` up to 
(but not including) `endtime` will be downloaded. Scan files are stored in and
read from the directory specified by `setSatelliteDataDir()`.

Downloading a single scan file:

```{r download_single_scan}
# Load library and set the scan file folder
library(MazamaSatelliteUtils)
setSatelliteDataDir("~/Data/Satellite")

scanFile <- goesaodc_downloadScanFiles(
  satID = "G17",
  datetime = "2020-09-08 12:30",
  timezone = "America/Los_Angeles"
)

scanFile

scanFile %>%
  goesaodc_convertFilenameToDatetime() %>%
  MazamaCoreUtils::parseDatetime(timezone = "America/Los_Angeles")
```

Downloading a series of scan files:

```{r download_scan_series}
scanFiles <- goesaodc_downloadScanFiles(
  satID = "G17",
  datetime = "2020-09-08 12:00",
  endtime = "2020-09-08 13:00",
  timezone = "America/Los_Angeles"
)

scanFiles

scanFiles %>%
  goesaodc_convertFilenameToDatetime() %>%
  MazamaCoreUtils::parseDatetime(timezone = "America/Los_Angeles")
```

You can also see what scan files are available locally or remotely with 
`goesaodc_listScanFiles()`:

```{r list_local_scan}
goesaodc_listScanFiles(
  satID = "G17",
  datetime = "2020-09-08 12:30",
  timezone = "America/Los_Angeles"
)
```

### 2. Creating Points and Rasters

Now that we can identify and download the scan files we're interested in, the
next step is to read them into R data structures. Here you have the choice of 
turning scan files into SpatialPointsDataFrames (SPDFs) with 
`goesaodc_createScanSpdf()` or rasters with `goesaodc_createScanRaster()`. 
These functions allow you to filter AOD readings using `bbox` and `dqfLevel`
parameters so that only readings within `bbox` will be kept and all readings 
with a DQF above `dqfLevel` will have their AOD values set to `NA`.

An important point to note: A raster is created by first generating a 
SpatialPointsDataFrame--which includes the filtering process--and *then* 
rasterizing the resulting points. A raster cell's AOD value is determined by 
summarizing the values of all the points that fall in it with the `fun` 
function (defaults to `mean`).

One more thing! Creating points and rasters requires that each scan reading is 
assigned to the proper place in its GOES grid. Therefore, before running either 
`goesaodc_createScan...()` function you must first install the GOES grid files 
with `goesaodc_installGoesGrids()`.

```{r install_goes_grids}
goesaodc_installGoesGrids()
```

Creating an SPDF:

```{r create_scan_spdf}
bboxOregon <- c(-125, -116, 42, 46.5)

scanSpdf <- goesaodc_createScanSpdf(
  filename = scanFile,
  bbox = bboxOregon,
  dqfLevel = 2
)
scanSpdf
```

Creating a raster:

```{r create_scan_raster}
scanRaster <- goesaodc_createScanRaster(
  filename = scanFile,
  bbox = bboxOregon,
  dqfLevel = 2,
  cellSize = 0.05
)
scanRaster
```

If you pass in a vector of filenames to `filename`, then 
`goesaodc_createScanSpdf()` will return a list of SPDFs and 
`goesaodc_createScanRaster()` will return a RasterBrick.

```{r create_scan_series_spdfs_and_rasterbrick}
scanSpdfList <- goesaodc_createScanSpdf(
  filename = scanFiles,
  bbox = bboxOregon
)

scanRasterBrick <- goesaodc_createScanRaster(
  filename = scanFiles,
  bbox = bboxOregon,
  cellSize = 0.05
)
```

An SPDF list or a RasterBrick can be used to aggregate AOD readings from
multiple scans into a single SPDF or RasterLayer. For instance, 
`goesaodc_calcAverageScan...()` will find the average value for each point/cell
across all the given scans:

```{r}
averageScanSpdf <- goesaodc_calcAverageScanSpdf(
  spdfList = scanSpdfList,
  na.rm = TRUE
)

averageScanRaster <- goesaodc_calcAverageScanRaster(
  rasterBrick = scanRasterBrick,
  na.rm = TRUE
)

```

The `goesaodc_calcTrendScan...()` functions calculate the "trend" for each 
point/cell across all the given scans. A trend value for a point/cell is
calculated by taking the difference between it's average value in the first half
of the scan list and it's average value in the second half.

```{r calc_trend_scan_points_and_rasters}
trendScanSpdf <- goesaodc_calcTrendScanSpdf(
  spdfList = scanSpdfList,
  na.rm = TRUE
)

trendScanRaster <- goesaodc_calcTrendScanRaster(
  rasterBrick = scanRasterBrick,
  na.rm = TRUE
)
```

Note: Rasters calculate their averages and trends using their cell values--not
the raw AOD values of the points they were made from.

### 3. Plotting

At last, we can draw SPDFs/rasters with `goesaodc_plotScanSpdf()` or
`goesaodc_plotScanRaster()`. These require you to pass in a single SPDF or 
RasterLayer respectively. Point and cell colors have a gradient color scale by
default, but you can set a discrete palette by passing in an RColorBrewer 
`paletteName` and a vector of `paletteBreaks` values. `NA` AOD values are drawn
in gray.

Plotting scan points:

```{r plot_scan_spdf, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
# Load state polygons
library(MazamaSpatialUtils)
setSpatialDataDir("~/Data/Spatial")
loadSpatialData("NaturalEarthAdm1")

goesaodc_plotScanSpdf(
  spdf = scanSpdf,
  bbox = bboxOregon,
  pointAlpha = 0.5,
  includeMap = TRUE,
  zoom = 7,
  stateCodes = "OR"
)
```

Plotting a scan raster:

```{r plot_scan_raster, fig.width=7, fig.height=5, fig.align="center"}
goesaodc_plotScanRaster(
  raster = scanRaster,
  bbox = bboxOregon,
  paletteName = "Blues",
  paletteBreaks = c(-Inf, 0, 1, 2, 3, 4, 5, Inf),
  stateCodes = "OR"
)
```

Since we have already seen how to aggregate multiple scans into a single SPDF or
RasterLayer, we can also draw averaged or trend scans:

```{r plot_aggregate_spdfs, fig.width=7, fig.height=5, fig.align="center"}
goesaodc_plotScanSpdf(
  spdf = averageScanSpdf,
  bbox = bboxOregon,
  stateCodes = "OR"
)

goesaodc_plotScanSpdf(
  spdf = trendScanSpdf,
  bbox = bboxOregon,
  stateCodes = "OR",
)
```

## Animations

`goesaodc_plotScanSpdf()` and `goesaodc_plotScanRaster()` allow you to plot 
static images of GOES AOD data. You can create animated videos of AOD as well 
using the `goesaodc_animateScanSpdfs_exec.R` and 
`goesaodc_animateScanRasters_exec.R` executable scripts. These scripts take in 
the same parameters as their plot counterparts (except `datetime` is now 
`starttime`), but they require several additional arguments:

* `satelliteDataDir`: The directory where satellite data is downloaded to and 
read from.
* `spatialDataDir`: The directory where spatial data is downloaded to and read 
from.
* `outputDir`: The directory where the video should be saved to.
* `logDir`: The directory where the output log files should be saved to.
* `frameRate`: The framerate of the video in fps.
* `verbose`: A logical flag whether to print out progress messages in the 
console.
* `version`: A logical flag whether to print out the current version of the 
executable script.

Another thing to note: Since animations are made up of individual plots, it's
essential that they all share a consistent color legend so that the color scale 
bounds don't change between frames and that the legend should still be drawn 
even if all the AOD values for a frame are `NA`. This is achieved by _requiring_
the user to set the `legendLimits` parameter.

Here's an example of animating points:

```
./goesaodc_animateScanSpdfs_exec.R \
  --satID="G17" \
  --starttime="2020-09-08 09:00" \
  --endtime="2020-09-08 10:00" \
  --timezone="America/Los_Angeles" \
  --bbox="-125, -116, 42, 47" \
  --dqfLevel=3 \
  --pointSize=0.3 \
  --pointShape=15 \
  --pointAlpha=0.6 \
  --paletteName="YlOrRd" \
  --legendLimits="-0.5, 5.5" \
  --includeMap=TRUE \
  --zoom=7 \
  --stateCodes="OR" \
  --satelliteDataDir="~/Data/Satellite" \
  --spatialDataDir="~/Data/Spatial" \
  --frameRate=6 \
  --outputDir="~/Downloads" \
  --logDir="~/Downloads" \
  --verbose=TRUE
```

And another for animating rasters:

```
./goesaodc_animateScanRasters_exec.R \
  --satID="G17" \
  --starttime="2020-09-08 09:00" \
  --endtime="2020-09-08 10:00" \
  --timezone="America/Los_Angeles" \
  --bbox="-125, -116, 42, 47" \
  --dqfLevel=3 \
  --cellSize=0.05 \
  --rasterAlpha=0.6 \
  --paletteName="YlOrRd" \
  --legendLimits="-0.5, 5.5" \
  --includeMap=TRUE \
  --zoom=7 \
  --stateCodes="OR" \
  --satelliteDataDir="~/Data/Satellite" \
  --spatialDataDir="~/Data/Spatial" \
  --frameRate=6 \
  --outputDir="~/Downloads" \
  --logDir="~/Downloads" \
  --verbose=TRUE
```
